### 2022-04-01 README.Rmd
intial 2

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Madrid zonas de bajas emisiones (low emission zones)

The repository is home to the analysis of low emissions zones in Madrid using spatial availability measure.

Currently, the following raw data from Madrid is contained here:

-  "parque_vehiculos_2021"  includes vehicle fleet characteristics of Madrid City. Of interest: the district and the environmental label associated to each vehicle.

- several excel documents from the Home Mobility Survey of the Community of Madrid. These documents include information like citizens' socioeconomic characteristics, origin and destination of the trip, transport mode, distance, time required, activities carried out at origin and destination, etc.

- A figure of Madrid spatial context.

<!-- badges: start -->
<!-- badges: end -->