### 2022-04-02 README.Rmd
Added all raw data and began a data preparation .rmd.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Madrid zonas de bajas emisiones (low emission zones)

The repository is home to the analysis of low emissions zones in Madrid using spatial availability measure.

Currently, the data contained in this repository is from the [Madrid City Council](https://datos.madrid.es/portal/site/egob/menuitem.c05c1f754a33a9fbe4b2e4b284f1a5a0/?vgnextoid=128b5dda42c84410VgnVCM2000000c205a0aRCRD&vgnextchannel=374512b9ace9f310VgnVCM100000171f5a0aRCRD&vgnextfmt=default) and the [Community of Madrid](https://datos.comunidad.madrid/catalogo/dataset/resultados-edm2018) databases retrieved March 31st 2022. It includes the following:

-  "parque_vehiculos_2021", "2020", and "2019",  includes vehicle fleet characteristics of Madrid City. Of interest: the district and the environmental label associated to each vehicle.

- several excel documents from the Home Mobility Survey of the Community of Madrid. These documents include information like citizens' socioeconomic characteristics, origin and destination of the trip, transport mode, distance, time required, activities carried out at origin and destination, etc.

- A figure of Madrid spatial context.

<!-- badges: start -->
<!-- badges: end -->