### 2022-04-22 00-OD-preparation.Rmd
Added visuals for Madrid Region and LEZ analysis in 00-OD-Prep.

---
title: "00-OD-preparation"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

Here we load the libraries we will be using in this .Rmd file:
```{r}
library(tidyverse) 
library(readxl) #imports excel files
library(geojsonio) #converts geoJSON files to R classes and vice versa
library(sf) # sf= simple features! These are spatial objects in R, read here: https://r-spatial.github.io/sf/index.html
library(lubridate) #for time formating
library(ggspatial)
```

## Import data

```{r}
OD_v <- read_excel("data-raw/EDM2018VIAJES.xlsx") #here we import the trips file, by default this function imports the first sheet in an excel file. This .Rmd file is saved in the same place as the .xlsx file so we just need to write the name of the file we want to read. run ' help(read_excel) ` to see the documentation of this function.

summary(OD_v) #descriptive stats of each row
```

You can also view a few rows of the 'OD' table like this:
```{r}
head(OD_v, 12) #the first 12 rows
```

Or view it in RStudio:
```{r}
#view(OD_v)
```

I'm interested in making the table spatial! It currently does not have coordinates but we can add from the Madrid Region shapefile which can be downloaded from the [Open Data Portal Madrid Regional Transport Consortium](https://datos.crtm.es/datasets/zonificacionzt1259/explore?location=40.420959%2C-3.695073%2C12.93):
```{r}
# # here we download the madrid zona shapefile using the API GeoJSON resource! neat
# zn <- geojson_read("https://opendata.arcgis.com/datasets/97f83bab03664d4e9853acf0e431d893_0.geojson",  what = "sp")
# zn<- st_as_sf(zn) %>% st_transform(crs = 4326) #transform it into a 'sf' class object and transform coordinates to WGS84
# 
# #let's save it to data-raw just in case this link/resource doesn't work in the future
# st_write(zn, "data-raw/zn.shp")

zn <- st_read("data-raw/zn.shp")
```

Importing a list of all the municipalities within the Madrid region and adding them to the zone shape file:
```{r}
t <- read_excel("data-raw/ZonificacionZT1259 complete.xlsx") %>% select(ZT1259, Municipality)

zn <- zn %>% merge(t, by = c("ZT1259"), all.x=TRUE)
```

Next, we will add a column which indicates if a zone is in a LEZ or is located within Madrid City Centre (this area contains the LEZ and is contained within the Municipality of Madrid):
```{r}
#IDS that correspond to CC - city center and LEZ - low emission zone
CC_IDS <- c("079-01", "079-02", "079-03", "079-04", "079-05", "079-06", "079-07")
LEZ_IDS <- c("079-01-001", "079-01-002", "079-01-003", "079-01-004", "079-01-005", "079-01-006", "079-01-007", "079-01-008", "079-01-009", "079-01-010", "079-01-012", "079-01-013", "079-01-014", "079-01-015", "079-01-016", "079-01-017", "079-01-018", "079-01-019","079-01-020", "079-01-021", "079-01-022", "079-01-023")

#manually added column to zn which identifies if an ID is in the city center (1) or not (0). This searches for any instance of "079-01-XXX", "079-02-XXX" etc.
zn$CC <- ifelse(grepl(paste(CC_IDS, collapse = "|"), zn$ZT1259), 1,0)

#more simple to use the mutate function here since the string of IDs is complete for LEZ
zn <- zn %>% mutate(LEZ = ifelse(ZT1259 %in% LEZ_IDS, 1, 0))

```

create a bounding box for the plot:
```{r}
#create the bounding box for the Toronto region + a little extra
bbox_new1 <- st_bbox(zn %>% filter(CC == 1)) # current bounding box

xrange <- bbox_new1$xmax - bbox_new1$xmin # range of x values
yrange <- bbox_new1$ymax - bbox_new1$ymin # range of y values

bbox_new1[1] <- bbox_new1[1] - (0.2 * xrange) # xmin - left
bbox_new1[3] <- bbox_new1[3] + (0.2 * xrange) # xmax - right
bbox_new1[2] <- bbox_new1[2] - (0.05 * yrange) # ymin - bottom
bbox_new1[4] <- bbox_new1[4] + (0.05 * yrange) # ymax - top

bbox_CityCentre <- bbox_new1 %>%  # take the bounding box ...
  st_as_sfc()
```

plot:
```{r}
plot_rmad <- ggplot() +
  geom_sf(data = zn, fill = "light grey", col = "dark grey", size=0.5) + #all of madrid region
  geom_sf(data = zn %>% filter(Municipality == "Madrid"), aes(fill = "Municipality of Madrid"), col = "yellow3", size=0.5, show.legend = "polygon") + # madrid municiality
  geom_sf(data = zn %>% filter(CC == 1), aes(fill =  "City centre"), col = NA, show.legend = "polygon") + # madrid city centre
  geom_sf(data = zn %>%  filter(LEZ == 1), aes(fill = "Low emission zone"), col = NA, show.legend = "polygon") + #madrid low emission zone
  geom_sf(data = bbox_CityCentre, fill = NA, col = "black", size = 0.5)+
  scale_fill_manual(labels = c("City centre", "Low emission zone", "Municipality of Madrid"),
                      values = c("green", "red", "yellow")) + 
  annotation_north_arrow(location = "tl", # north arrow for both the main plot and inset
                         height = unit(0.8, "cm"), 
                         width = unit(0.8, "cm"),
                         style = north_arrow_orienteering(line_width = 0.25,
                                                          line_col = "dimgrey", 
                                                          fill = c("white","dimgrey"))) +
  annotation_scale(bar_cols = c("dimgrey", "white"), # scale bar for both the main plot and inset
                   height = unit(0.15, "cm")) +
  scale_y_continuous(breaks = c(40, 40.4, 40.8, 41.2)) + 
  scale_x_continuous(breaks = c(-3.0, -3.8, -4.6)) +
  guides(fill=guide_legend(title=NULL)) +
  theme_minimal() + theme(legend.position = c(0.25, .16))

#just the city center and LEZ
plot_ccmad <- ggplot() +
  geom_sf(data = zn %>% filter(CC == 1), aes(fill = "City centre"), col = "green3") + # madrid city centre
  geom_sf(data = zn %>%  filter(LEZ == 1), aes(fill = "Low emission zone"), col = "red4") + #madrid low emission zone
  geom_sf(data = bbox_CityCentre, fill = NA, col = "black", size = 0.5)+
  scale_fill_manual(labels = c("City centre", "Low emission zone"),
                    values = c("green", "red")) + 
  annotation_scale(bar_cols = c("dimgrey", "white"), # scale bar for both the main plot and inset
                   height = unit(0.15, "cm"),
                   pad_y = unit(0.05, "cm")) +
  guides(fill=guide_legend(title=NULL)) +
  theme_void() + theme(legend.position = "none")

cowplot::plot_grid(plot_rmad, plot_ccmad, ncol = 2, rel_widths = c(2.5,1)) 
```

## Organizing the trip level data

```{r}
#adding travel time... we don't even need r5r! we just need to decide how to remove A polluting cars
OD_v <- OD_v %>% mutate(
  TIEMPO_VIAJES = 
                  as.numeric((as.POSIXct(OD_v$VDESHORAFIN, format="%H%M") -
                  as.POSIXct(OD_v$VORIHORAINI, format="%H%M"))/60))

summary(OD_v$TIEMPO_VIAJES)

```

## Exploring the data - filtering OD trips, home to work

```{r}
OD_v_Oc_Dt <- OD_v %>% filter(VORI == "1" & VDES == "2") #only filter in origin casa trips and destination trabajo

count(OD_v_Oc_Dt) # from the 222744 trips we now have 27,796 trips
```

What can we say about these work trips?

Well, the majority of trip frequency is "A diario, de lunes a viernes" (Daily, from Monday to Friday - 1)
```{r}
hist(OD_v_Oc_Dt$VFRECUENCIA)
```
The average destination arrival time (Destino. Hora llegada) is:
```{r}
Dtrips <- as.POSIXct(OD_v$VDESHORAFIN, format="%H%M") %>% format("%H:%M") # formatting and removing date so we just how time
seconds_to_period(mean(period_to_seconds(hm(Dtrips)), na.rm=T)) # finding mean
```
9:48am arrival.. not very helpful.

Let's see the distribution:
```{r}
hist(OD_v_Oc_Dt$VDESHORAFIN%>% as.numeric(), breaks = 24)
```
So we see that 8AM, and 9AM arrival times are most frequent.

We also know that the majority do make this trip by personal car "¿Dispone de vehículo particular para hacer este viaje?" (1):
```{r}
hist(OD_v_Oc_Dt$VVEHICULO)

OD_v_Oc_Dt$VVEHICULO %>% typeof()
```

Let's see how long are short-trips are (by distance, assuming in km units):
```{r}
t <- OD_v_Oc_Dt %>% filter(DISTANCIA_VIAJE <= 60) %>% select(DISTANCIA_VIAJE) %>% unlist() %>% as.numeric() 
hist(t)
```

As you can see though there are some long distance trip (greater than 60km). There are 101 trips which are between 60km to 600km and 1 trip greater than 600km. For perspective though, these trips make up onl 0.3% of total trips taken ( ~100 trips out of ~28,000 trips).
```{r}
t <- OD_v_Oc_Dt %>% filter(DISTANCIA_VIAJE >= 60 & DISTANCIA_VIAJE <= 600) %>% select(DISTANCIA_VIAJE) %>% unlist() %>% as.numeric() 
hist(t, breaks = 50)

t <- OD_v_Oc_Dt %>% filter(DISTANCIA_VIAJE >= 600) %>% select(DISTANCIA_VIAJE) %>% unlist() %>% as.numeric() 
hist(t, breaks = 50)
```

Let's check the primary mode: 
```{r}
hist(OD_v_Oc_Dt$MODO_PRIORITARIO)
```

MODO_PRIORITARIO -> 11, 12, 13, 14, 15, 16 correspond to:
11 '11. Private car - driver
12 '12. Company car - driver
13 '13. Rental car without driver
14 '14. Private car - passenger
15 '15. Company car- passenger
16 '16. Car rental with driver - passenger

## Making OD home to work trips spatial

Here we add # of origin trips to the zona sf object, so we see where trips are coming from.

First, we count the number of trips coming to each destination:
```{r}
Dtrips_zona <- OD_v_Oc_Dt %>% group_by(VDESZT1259)  %>% summarise(Dtrips = (n())) # we group by the Destination ZT1259 ID and then we summarize the number of rows that have the same ZT using the summarize() function and the n() function to create a new 'Dtrips' column.

head(Dtrips_zona,12)
```

Now, we add this new column to the zonas sf object:
```{r}
zn_t <- zn %>% merge(Dtrips_zona, by.x = "ZT1259", by.y="VDESZT1259")

head(zn_t %>% st_drop_geometry(), 12) # notice the new 'Otrips' column
```

a simple plot of the number of DESTINATION home-based trips:
```{r}
ggplot() +
  geom_sf(data = zn_t,
          aes(fill = Dtrips),
          color  = NA) +
  scale_fill_distiller(palette = "Spectral")
```

For comparison... lets plot the number of ORIGIN home-based trips:
```{r}
Otrips_zona <- OD_v_Oc_Dt %>% group_by(VORIZT1259)  %>% summarise(Otrips = (n())) # we group by the origin ZT1259 ID and then we summarize the number of rows that have the same ZT using the summarize() function and the n() function to create a new 'Otrips' column.

zn_t <- zn_t %>% merge(Otrips_zona, by.x = "ZT1259", by.y="VORIZT1259", all.x = TRUE)

ggplot() +
  geom_sf(data = zn_t,
          aes(fill = Otrips),
          color  = NA) +
  scale_fill_distiller(palette = "Spectral")
```
How many trips come into the LEZ from outside of the LEZ and how many of those are car trips? Just curious.

First, let's add a column which correspond to if the trip is made by car or not my car.
```{r}
OD_v_Oc_Dt <- OD_v_Oc_Dt %>% mutate(bycar = ifelse(MODO_PRIORITARIO %in% c(11, 12, 13, 14, 15, 16),
                                                                         1, 0),
                                    nocar = ifelse(!(MODO_PRIORITARIO %in% c(11, 12, 13, 14, 15, 16)),
                                                                         1, 0))
```

Next, let's split the data - one subset of all trips that don't go into the LEZ, and another subset of trips that DO go into the LEZ
```{r}
OD_v_Oc_Dt_noDesLEZ <- OD_v_Oc_Dt %>% filter(!(VDESZT1259 %in% LEZ_IDS) | (VORIZT1259 %in% LEZ_IDS))
OD_v_Oc_Dt_DesLEZ <- OD_v_Oc_Dt %>% filter((VDESZT1259 %in% LEZ_IDS) & !(VORIZT1259 %in% LEZ_IDS))

count(OD_v_Oc_Dt_DesLEZ)/count(OD_v_Oc_Dt)*100
count(OD_v_Oc_Dt_noDesLEZ) + count(OD_v_Oc_Dt_DesLEZ) ==  count(OD_v_Oc_Dt)
```
we can see that trips into LEZ are a small proportion - only 5.84% of total home-to-work trips in the madrid region have a destination in the LEZ (and do not originate in the LEZ). Now... how many of these are car trips?

```{r}
sum(OD_v_Oc_Dt_DesLEZ$bycar)/count(OD_v_Oc_Dt_DesLEZ)
```
only 19% of these LEZ trips are made by car! This may not have a huge GHG emission savings... but perhaps it has a big impact on equity?

Let's plot where these car LEZ trips are coming from:
```{r}
D2trips_zona <- OD_v_Oc_Dt_DesLEZ %>% group_by(VORIZT1259)  %>% summarise(Dtrips = (n()),
                                                                              bycar = sum(bycar),
                                                                              nocar = sum(nocar))

zn_t <- zn %>% merge(D2trips_zona, by.x = "ZT1259", by.y="VORIZT1259")
```

plots:
```{r}
# % of trips made by car
ggplot() +
  geom_sf(data = zn_t,
          aes(fill = (bycar/Dtrips * 100) ),
          color  = NA) +
  scale_fill_distiller(palette = "Spectral") +
  labs(title = "% work trips by origin (excluding those that originate in LEZ) made by car ")

# magnitude of trips made by car 
ggplot() +
  geom_sf(data = zn_t,
          aes(fill = (bycar) ),
          color  = NA) +
  scale_fill_distiller(palette = "Spectral") +
  labs(title = "number of work trips by origin (excluding those that originate in LEZ) made by car ")
```

% of car trips which go into the LEZ 
```{r}
# Mag
ggplot() +
  geom_sf(data = zn_t,
          aes(fill = (bycar/Dtrips * 100) ),
          color  = NA) +
  scale_fill_distiller(palette = "Spectral") +
  labs(title = "% work trips (excluding those that originate in LEZ) made by car ")
```

```{r}
ggplot() +
  geom_sf(data = zn %>% filter(CC == 1), aes(fill = "City centre"), col = "green3") + # madrid city centre
  geom_sf(data = zn %>%  filter(LEZ == 1), aes(fill = "Low emission zone"), col = "red4") + #madrid low emission zone
  geom_sf(data = bbox_CityCentre, fill = NA, col = "black", size = 0.5)+
  scale_fill_manual(labels = c("City centre", "Low emission zone"),
                    values = c("green", "red")) + 
  annotation_scale(bar_cols = c("dimgrey", "white"), # scale bar for both the main plot and inset
                   height = unit(0.15, "cm"),
                   pad_y = unit(0.05, "cm")) +
  guides(fill=guide_legend(title=NULL)) +
  theme_void() + theme(legend.position = "none"
```

## Saving data
```{r}
save(zn, file = "data/zn.RData")
save(OD_v, file = "data/OD_v.RData")
```