### 2022-05-08 Draft-manuscript-tobemoved.Rmd
finalized 2 spatial availability scenario calculations (baseline no LEZ and optimistic LEZ)

---
title: "Draft-manuscript-tobemoved"
output: html_document
---

```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = '', 
  out.width = "1\\linewidth")
```

```{r install-data-package, eval = FALSE, include=FALSE}
# if (!require("MadridZBE", character.only = TRUE)) {
#       remotes::install_github("soukhova/MadridZBE",
#                         build_vignettes = TRUE)
#   }
```

```{r load-packages, include=FALSE, cache=FALSE}
library(tidyverse)
library(sf)
library(fitdistrplus)
library(ggplot2)
library(readxl) #imports excel files
library(patchwork)
library(cowplot)
```

```{r sp_avail-function,include=FALSE}
#defining the spatial availability function
sp_avail <- function(x, o_id, d_id, pop, opp, r, f, alpha = 1){

  o_id <- rlang::enquo(o_id)
  d_id <- rlang::enquo(d_id)
  pop <- rlang::enquo(pop)
  opp <- rlang::enquo(opp)
  r <- rlang::enquo(r)
  f <- rlang::enquo(f)

  sum_pop <- x %>%
    dplyr::distinct(!!o_id,
                    .keep_all = TRUE) %>%
    dplyr::mutate(sum_pop = !!r*(!!pop)^alpha) %>%
    dplyr::pull(sum_pop) %>%
    sum()

  f_p <- dplyr::pull(x, !!r) * dplyr::pull(x, !!pop)^alpha / sum_pop

  sum_impedance <- x %>%
    dplyr::group_by(!!d_id) %>%
    dplyr::summarize(sum_impedance = sum(!!f))

  x <- x %>%
    dplyr::left_join(sum_impedance,
                     by = rlang::as_name(d_id))

  f_c <- dplyr::pull(x, !!f) / x$sum_impedance

  x$f_c <- f_c
  x$f_p <- f_p

  sum_pa <- x %>%
    dplyr::group_by(!!d_id) %>%
    dplyr::summarize(sum_pa= sum(f_p * f_c))

  x <- x %>%
    dplyr::left_join(sum_pa,
                     by = rlang::as_name(d_id))
  f_t <- (f_p * f_c) / dplyr::pull(x, sum_pa)

  dplyr::pull(x, !!opp) * f_t
}
```

\newpage

# Introduction

- Some information about LEZ generally

- Information about accessibility issues regarding LEZ; the tradition of using accessibility to evaluate these differences. Mention of spatial availability.

- What we will be doing, the aim.

\newpage

# Using accessibility to evaluate inequity background

- Information about inequity and LEZ may exasperate these issues

- lit has found that emissions reduciton should not result in inequalities..

\newpage

# Methods

## Data 
- -plots of madrid zone, description of the LEZ scheme, how effective it is now and will be in the future - change in modal shares and emissions but what about equity

## Estimated data
- prediction in change in travel mode
- prediction in private vehicle fleet label A/B/C proportions

## Spatial availability
- spatial availability explanation 
  - impedance function explanation

\newpage

# Results

## Fleet characteristics today vs. 2025
<!-- to do -->

## Impedance function calibration

```{r data-import-filter}
OD_v <- read_excel("data-raw/EDM2018VIAJES.xlsx")

OD_v_HW <- OD_v %>% filter(VORI == "1" & VDES == "2") #only filter in origin casa trips and destination trabajo

OD_v_HW <- OD_v_HW %>% 
  mutate(TIEMPO_VIAJES = 
                  as.numeric((as.POSIXct(OD_v_HW$VDESHORAFIN, format="%H%M") -
                  as.POSIXct(OD_v_HW$VORIHORAINI, format="%H%M"))/60)) %>%
  mutate(TIEMPO_VIAJES = ifelse(VDESHORAFIN == "2410", 40, 
                                         ifelse(VDESHORAFIN == "2440", 55, 
                                                ifelse(VDESHORAFIN == "2800", 420, TIEMPO_VIAJES)))) %>% # some destination times are greater than 2400 and so yield an NA different, correct these 3 trips
  mutate(TIEMPO_VIAJES = ifelse(TIEMPO_VIAJES == 0, 0.1, TIEMPO_VIAJES)) # set all 0min travel times to 0.1 min

summary(OD_v_HW$TIEMPO_VIAJES)
```

```{r data-for-impedance}
#here we separate all the "car" trips, "transit" trips, "bike" trips, and "walk" trips. We then will create 4 different imepedance function....
transit_HW_alltt <- OD_v_HW  %>% filter(MODO_PRIORITARIO <= 9) %>%
  dplyr::select(TIEMPO_VIAJES)

car_motor_HW_alltt <- OD_v_HW  %>% filter(MODO_PRIORITARIO <= 19 & MODO_PRIORITARIO >= 10) %>%
  dplyr::select(TIEMPO_VIAJES) #includes both motorcycles and cars

bike_HW_alltt <- OD_v_HW  %>% filter(MODO_PRIORITARIO <= 23 & MODO_PRIORITARIO >= 20) %>%
  dplyr::select(TIEMPO_VIAJES)

walk_HW_alltt <- OD_v_HW  %>% filter(MODO_PRIORITARIO == 24) %>%
  dplyr::select(TIEMPO_VIAJES)

```

plotting histograms:
```{r}
hist(transit_HW_alltt%>% unlist() %>% as.numeric(), breaks=100)
hist(car_motor_HW_alltt%>% unlist() %>% as.numeric(), breaks=100)
hist(bike_HW_alltt%>% unlist() %>% as.numeric(), breaks=100)
hist(walk_HW_alltt%>% unlist() %>% as.numeric(), breaks=100)
```
```{r}
#cullen grey graph - BUS
descdist(transit_HW_alltt%>% unlist() %>% as.numeric(), discrete=FALSE, boot=500)
```

```{r bus-itting-impedance-function}
# using fitdist function to fit a distribution using the default maximum likelihood estimation method and Nelder-Mead method for direct optimization
gamma_ <- fitdistrplus::fitdist(data=transit_HW_alltt%>% unlist() %>% as.numeric(), "gamma", method="mle", optim.method="Nelder-Mead") 
lnorm_ <- fitdistrplus::fitdist(data=transit_HW_alltt%>% unlist() %>% as.numeric(), "lnorm", method="mle", optim.method="Nelder-Mead")
norm_ <-fitdistrplus::fitdist(data=transit_HW_alltt%>% unlist() %>% as.numeric(), "norm", method="mle", optim.method="Nelder-Mead")
exp_ <- fitdistrplus::fitdist(data=transit_HW_alltt%>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead")
# pois_ <- fitdistrplus::fitdist(data=transit_HW_alltt%>% unlist() %>% as.numeric(), "pois", method="mle", optim.method="Nelder-Mead")
# nbinom_ <- fitdistrplus::fitdist(data=transit_HW_alltt%>% unlist() %>% as.numeric(), "nbinom", method="mle", optim.method="Nelder-Mead")
# geom_ <- fitdistrplus::fitdist(data=transit_HW_alltt%>% unlist() %>% as.numeric(), "geom", method="mle", optim.method="Nelder-Mead")
# beta_ <- fitdistrplus::fitdist(data=transit_HW_alltt%>% unlist() %>% as.numeric(), "beta", method="mle", optim.method="Nelder-Mead")
logis_ <- fitdistrplus::fitdist(data=transit_HW_alltt%>% unlist() %>% as.numeric(), "logis", method="mle", optim.method="Nelder-Mead")

plot(gamma_)
plot(lnorm_)
plot(norm_)
plot(exp_)
# plot(pois_)
# plot(nbinom_)
# plot(geom_)
# plot(beta_)
plot(logis_)
```
now let's compare these models AIC and BIC to see which fits the bus data best:
```{r}
broom::glance(MASS::fitdistr(transit_HW_alltt%>% unlist() %>% as.numeric(),"gamma"))
broom::glance(MASS::fitdistr(transit_HW_alltt%>% unlist() %>% as.numeric(),"lognormal"))
broom::glance(MASS::fitdistr(transit_HW_alltt%>% unlist() %>% as.numeric(),"normal"))
broom::glance(MASS::fitdistr(transit_HW_alltt%>% unlist() %>% as.numeric(),"exponential"))
# plot(pois_)
# plot(nbinom_)
# plot(geom_)
# plot(beta_)
broom::glance(MASS::fitdistr(transit_HW_alltt%>% unlist() %>% as.numeric(),"logistic"))
```

lognormal has largest logLik and the smallest AIC and BIC. we will pick this lognormal function for transit trips.
```{r}
HW_transit_lnorm <- lnorm_
```

Let's do car/motorcycles now:
```{r car-motor-itting-impedance-function}
# using fitdist function to fit a distribution using the default maximum likelihood estimation method and Nelder-Mead method for direct optimization
gamma_ <- fitdistrplus::fitdist(data=car_motor_HW_alltt%>% unlist() %>% as.numeric(), "gamma", method="mle", optim.method="Nelder-Mead") 
lnorm_ <- fitdistrplus::fitdist(data=car_motor_HW_alltt%>% unlist() %>% as.numeric(), "lnorm", method="mle", optim.method="Nelder-Mead")
norm_ <-fitdistrplus::fitdist(data=car_motor_HW_alltt%>% unlist() %>% as.numeric(), "norm", method="mle", optim.method="Nelder-Mead")
exp_ <- fitdistrplus::fitdist(data=car_motor_HW_alltt%>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead")
# pois_ <- fitdistrplus::fitdist(data=car_motor_HW_alltt%>% unlist() %>% as.numeric(), "pois", method="mle", optim.method="Nelder-Mead")
# nbinom_ <- fitdistrplus::fitdist(data=car_motor_HW_alltt%>% unlist() %>% as.numeric(), "nbinom", method="mle", optim.method="Nelder-Mead")
# geom_ <- fitdistrplus::fitdist(data=car_motor_HW_alltt%>% unlist() %>% as.numeric(), "geom", method="mle", optim.method="Nelder-Mead")
# beta_ <- fitdistrplus::fitdist(data=car_motor_HW_alltt%>% unlist() %>% as.numeric(), "beta", method="mle", optim.method="Nelder-Mead")
logis_ <- fitdistrplus::fitdist(data=car_motor_HW_alltt%>% unlist() %>% as.numeric(), "logis", method="mle", optim.method="Nelder-Mead")

plot(gamma_)
plot(lnorm_)
plot(norm_)
plot(exp_)
# plot(pois_)
# plot(nbinom_)
# plot(geom_)
# plot(beta_)
plot(logis_)
```
now let's compare these models AIC and BIC to see which fits the car data best:
```{r}
broom::glance(MASS::fitdistr(car_motor_HW_alltt%>% unlist() %>% as.numeric(),"gamma"))
broom::glance(MASS::fitdistr(car_motor_HW_alltt%>% unlist() %>% as.numeric(),"lognormal"))
broom::glance(MASS::fitdistr(car_motor_HW_alltt%>% unlist() %>% as.numeric(),"normal"))
broom::glance(MASS::fitdistr(car_motor_HW_alltt%>% unlist() %>% as.numeric(),"exponential"))
# plot(pois_)
# plot(nbinom_)
# plot(geom_)
# plot(beta_)
broom::glance(MASS::fitdistr(car_motor_HW_alltt%>% unlist() %>% as.numeric(),"logistic"))
```
gamma has largest logLik and the smallest AIC and BIC. we will pick this gamma function for car trips.
```{r}
HW_car_motor_gamma <- gamma_
```

Let's do bike now:
```{r bike-itting-impedance-function}
# using fitdist function to fit a distribution using the default maximum likelihood estimation method and Nelder-Mead method for direct optimization
gamma_ <- fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "gamma", method="mle", optim.method="Nelder-Mead") 
lnorm_ <- fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "lnorm", method="mle", optim.method="Nelder-Mead")
norm_ <-fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "norm", method="mle", optim.method="Nelder-Mead")
exp_ <- fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead")
# pois_ <- fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "pois", method="mle", optim.method="Nelder-Mead")
# nbinom_ <- fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "nbinom", method="mle", optim.method="Nelder-Mead")
# geom_ <- fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "geom", method="mle", optim.method="Nelder-Mead")
# beta_ <- fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "beta", method="mle", optim.method="Nelder-Mead")
logis_ <- fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "logis", method="mle", optim.method="Nelder-Mead")

plot(gamma_)
plot(lnorm_)
plot(norm_)
plot(exp_)
plot(logis_)
```
now let's compare these models AIC and BIC to see which fits the bike data best:
```{r}
broom::glance(MASS::fitdistr(bike_HW_alltt%>% unlist() %>% as.numeric(),"gamma"))
broom::glance(MASS::fitdistr(bike_HW_alltt%>% unlist() %>% as.numeric(),"lognormal"))
broom::glance(MASS::fitdistr(bike_HW_alltt%>% unlist() %>% as.numeric(),"normal"))
broom::glance(MASS::fitdistr(bike_HW_alltt%>% unlist() %>% as.numeric(),"exponential"))
broom::glance(MASS::fitdistr(bike_HW_alltt%>% unlist() %>% as.numeric(),"logistic"))
```
lognormal has largest logLik and the smallest AIC and BIC. we will pick lognormal function for bike trips! Interesting how car mode is different from all other modes.

```{r}
HW_bike_lnorm <- lnorm_
```

Let's do walk now:
```{r walk-itting-impedance-function}
# using fitdist function to fit a distribution using the default maximum likelihood estimation method and Nelder-Mead method for direct optimization
gamma_ <- fitdistrplus::fitdist(data=walk_HW_alltt%>% unlist() %>% as.numeric(), "gamma", method="mle", optim.method="Nelder-Mead") 
lnorm_ <- fitdistrplus::fitdist(data=walk_HW_alltt%>% unlist() %>% as.numeric(), "lnorm", method="mle", optim.method="Nelder-Mead")
norm_ <-fitdistrplus::fitdist(data=walk_HW_alltt%>% unlist() %>% as.numeric(), "norm", method="mle", optim.method="Nelder-Mead")
exp_ <- fitdistrplus::fitdist(data=walk_HW_alltt%>% unlist() %>% as.numeric(), "exp", method="mle", optim.method="Nelder-Mead")
# pois_ <- fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "pois", method="mle", optim.method="Nelder-Mead")
# nbinom_ <- fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "nbinom", method="mle", optim.method="Nelder-Mead")
# geom_ <- fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "geom", method="mle", optim.method="Nelder-Mead")
# beta_ <- fitdistrplus::fitdist(data=bike_HW_alltt%>% unlist() %>% as.numeric(), "beta", method="mle", optim.method="Nelder-Mead")
logis_ <- fitdistrplus::fitdist(data=walk_HW_alltt%>% unlist() %>% as.numeric(), "logis", method="mle", optim.method="Nelder-Mead")

plot(gamma_)
plot(lnorm_)
plot(norm_)
plot(exp_)
plot(logis_)
```
now let's compare these models AIC and BIC to see which fits the walk data best:
```{r}
broom::glance(MASS::fitdistr(walk_HW_alltt%>% unlist() %>% as.numeric(),"gamma"))
broom::glance(MASS::fitdistr(walk_HW_alltt%>% unlist() %>% as.numeric(),"lognormal"))
broom::glance(MASS::fitdistr(walk_HW_alltt%>% unlist() %>% as.numeric(),"normal"))
broom::glance(MASS::fitdistr(walk_HW_alltt%>% unlist() %>% as.numeric(),"exponential"))
broom::glance(MASS::fitdistr(walk_HW_alltt%>% unlist() %>% as.numeric(),"logistic"))
```
lognormal has largest logLik and the smallest AIC and BIC. we will pick lognormal function for walk trips!

```{r}
HW_walk_lnorm <- lnorm_
```

Add the number of trips (by mode) and create objects which contain the number of workers (origin) and jobs (destinations) - by mode:
```{r}
OD_v_HW <- OD_v_HW  %>% mutate(trips_transit = ifelse(MODO_PRIORITARIO <= 9, 1,0),
                    trips_car_motor = ifelse(MODO_PRIORITARIO <= 19 & MODO_PRIORITARIO >= 10, 1, 0),
                    trips_bike = ifelse(MODO_PRIORITARIO <= 23 & MODO_PRIORITARIO >= 20, 1, 0),
                    trips_walk = ifelse(MODO_PRIORITARIO == 24, 1, 0))

workers <- OD_v_HW %>%
  group_by(VORIZT1259) %>%
  summarize(workers_transit = sum(trips_transit),
            workers_car = sum(trips_car_motor),
            workers_bike = sum(trips_bike),
            workers_walk = sum(trips_walk),
            .groups = "drop") %>%
  transmute(ZT1259 = VORIZT1259,
            workers_transit,
            workers_car,
            workers_bike,
            workers_walk)

jobs <- OD_v_HW %>%
  group_by(VDESZT1259) %>%
  summarize(jobs_transit = sum(trips_transit),
            jobs_car = sum(trips_car_motor),
            jobs_bike = sum(trips_bike),
            jobs_walk = sum(trips_walk),
            .groups = "drop") %>%
  transmute(ZT1259 = VDESZT1259,
            jobs_transit,
            jobs_car,
            jobs_bike,
            jobs_walk)
```

Join workers and jobs to the ZN simple features object:
```{r}
load("data/zn.RData")

zn <- zn %>%
  # Join workers
  left_join(workers,
            by = "ZT1259") %>%
  # Join jobs
  left_join(jobs,
            by = "ZT1259") %>%
  # Replace NAs with zeros
  mutate(workers_transit = tidyr::replace_na(workers_transit, 0),
         workers_car = tidyr::replace_na(workers_car, 0),
         workers_bike = tidyr::replace_na(workers_bike, 0),
         workers_walk = tidyr::replace_na(workers_walk, 0),
         jobs_transit = tidyr::replace_na(jobs_transit, 0),
         jobs_car = tidyr::replace_na(jobs_car, 0),
         jobs_bike = tidyr::replace_na(jobs_bike, 0),
         jobs_walk = tidyr::replace_na(jobs_walk, 0))

# sum(zn$workers_transit) + sum(zn$workers_bike) + sum(zn$workers_car) + sum(zn$workers_walk)
# sum(zn$jobs_transit) + sum(zn$jobs_bike) + sum(zn$jobs_car) + sum(zn$jobs_walk)
```

add the number of workers and jobs from each zona to each trip:
```{r}
#add the number of jobs and workers to the od_ft_tt matrix
OD <- OD_v_HW %>% merge(zn %>% dplyr::select(ZT1259, workers_transit,workers_car,workers_bike,workers_walk) %>% st_drop_geometry(),
                   by.x = "VORIZT1259", by.y="ZT1259", all.x = TRUE)

OD <- OD %>% merge(zn %>% dplyr::select(ZT1259, jobs_transit,jobs_car,jobs_bike,jobs_walk) %>% st_drop_geometry(),
                   by.x = "VDESZT1259", by.y="ZT1259", all.x = TRUE)

OD <- OD %>% 
  mutate(workers_transit = ifelse(is.na(workers_transit), 0, workers_transit),
         workers_car = ifelse(is.na(workers_car), 0, workers_car),
         workers_bike = ifelse(is.na(workers_bike), 0, workers_bike),
         workers_walk = ifelse(is.na(workers_walk), 0, workers_walk),
         jobs_car = ifelse(is.na(jobs_car), 0, jobs_car),
         jobs_transit = ifelse(is.na(jobs_transit), 0, jobs_transit),
         jobs_bike = ifelse(is.na(jobs_bike), 0, jobs_bike),
         jobs_walk = ifelse(is.na(jobs_walk), 0, jobs_walk))
```

now reshape the data, each row will represent an any origin to to 1 destination (i.e., destination row is unique). We will average the travel times, by mode, and then use this average travel time to find the 'f' (impedance value for travel for each mode):
```{r}
#consolidate trips so each trip is unique. workers = jobs.
OD_test <- OD %>%
  group_by(VORIZT1259,VDESZT1259) %>%
  summarise(trips_transit = sum(trips_transit),
            trips_car_motor = sum(trips_car_motor),
            trips_bike = sum(trips_bike),
            trips_walk = sum(trips_walk),
            workers_transit = mean(workers_transit),
            # workers_transit = ifelse(is.na(workers_transit), 0, workers_transit),
            workers_car = mean(workers_car),
            # workers_car = ifelse(is.na(workers_car), 0, workers_car),
            workers_bike = mean(workers_bike),
            # workers_bike = ifelse(is.na(workers_bike), 0, workers_bike),
            workers_walk = mean(workers_walk),
            # workers_walk = ifelse(is.na(workers_walk), 0, workers_walk),
            jobs_transit = mean(jobs_transit),
            jobs_car = mean(jobs_car),
            jobs_bike = mean(jobs_bike),
            jobs_walk = mean(jobs_walk),
            # jobs_car = ifelse(is.na(jobs_car), 0, jobs_car),
            # jobs_transit = ifelse(is.na(jobs_transit), 0, jobs_transit),
            # jobs_bike = ifelse(is.na(jobs_bike), 0, jobs_bike),
            # jobs_walk = ifelse(is.na(jobs_walk), 0, jobs_walk),
            tt_transit = ifelse(trips_transit != 0, mean(TIEMPO_VIAJES, na.rm=T), NA),
            tt_car = ifelse(trips_car_motor != 0, mean(TIEMPO_VIAJES, na.rm=T), NA),
            tt_bike = ifelse(trips_bike != 0, mean(TIEMPO_VIAJES, na.rm=T), NA),
            tt_walk = ifelse(trips_walk != 0, mean(TIEMPO_VIAJES, na.rm=T), NA)) %>% ungroup()

# OD_test <- OD %>%
#   mutate(workers_ = (workers_car+ workers_transit+ workers_bike+ workers_walk),
#          workers_ = ifelse(is.na(workers_), 0, workers_),
#          workers_car = ifelse(is.na(workers_car), 0, workers_car),
#          workers_transit = ifelse(is.na(workers_transit), 0, workers_transit),
#          workers_bike = ifelse(is.na(workers_bike), 0, workers_bike),
#          workers_walk = ifelse(is.na(workers_walk), 0, workers_walk),
#            jobs_ = (jobs_car+ jobs_transit+ jobs_bike+ jobs_walk),
#          jobs_ = ifelse(is.na(jobs_), 0, jobs_),
#          jobs_car = ifelse(is.na(jobs_car), 0, jobs_car),
#          jobs_transit = ifelse(is.na(jobs_transit), 0, jobs_transit),
#          jobs_bike = ifelse(is.na(jobs_bike), 0, jobs_bike),
#          jobs_walk = ifelse(is.na(jobs_walk), 0, jobs_walk),
#          tt_transit = ifelse(trips_transit != 0, TIEMPO_VIAJES, NA),
#             tt_car = ifelse(trips_car_motor != 0, TIEMPO_VIAJES, NA),
#             tt_bike = ifelse(trips_bike != 0, TIEMPO_VIAJES, NA),
#             tt_walk = ifelse(trips_walk != 0, TIEMPO_VIAJES, NA)) %>% ungroup()

#check, should add to the origin number of trips 27795 - perfect!
# sum(OD$trips_transit) + sum(OD$trips_car_motor) + sum(OD$trips_bike) + sum(OD$trips_walk)
```

now estimate f value using these averaged travel times:
```{r, echo=FALSE, warning=FALSE, message=FALSE}
#dgamma is the density plot for gamma function and dlnorm is the desntiy plot for the lognormal function; adding values as "f" our impedance function
OD_test <- OD_test %>%
  mutate(f_car = dgamma(tt_car, HW_car_motor_gamma$estimate["shape"], HW_car_motor_gamma$estimate["rate"]),
         f_car = ifelse(is.na(f_car), 0, f_car),
         f_transit = dlnorm(tt_transit, HW_transit_lnorm$estimate["meanlog"], HW_transit_lnorm$estimate["sdlog"]),
         f_transit = ifelse(is.na(f_transit), 0, f_transit),
         f_bike = dlnorm(tt_bike, HW_bike_lnorm$estimate["meanlog"], HW_bike_lnorm$estimate["sdlog"]),
         f_bike = ifelse(is.na(f_bike), 0, f_bike),
         f_walk = dlnorm(tt_walk, HW_walk_lnorm$estimate["meanlog"], HW_walk_lnorm$estimate["sdlog"]),
         f_walk = ifelse(is.na(f_walk), 0, f_walk))
```

```{=tex}
\begin{equation}
\label{gamma-dist}
\begin{array}{l}\ 
f(x, \alpha, \beta) = \frac {x^{\alpha-1}e^{-\frac{x}{\beta}}}{ \beta^{\alpha}\Gamma(\alpha)} \quad \text{for } 0 \leq x \leq \infty\\

\Gamma(\alpha) =  \int_{0}^{\infty} x^{\alpha-1}e^{-x} \,dx\\
\end{array}
\end{equation}
```

## Spatial availability comparision

```{r}
#calculate spatial availability
OD_SA <- OD_test %>%
  mutate(catch = 1) %>%
  mutate(SA_ij_car = sp_avail(., 
                         o_id = VORIZT1259,
                         d_id = VDESZT1259,
                         pop = workers_car,
                         opp = jobs_car,
                         r = catch,
                         f = f_car),
         SA_ij_car = ifelse(is.nan(SA_ij_car), 0, SA_ij_car),
         SA_ij_transit = sp_avail(., 
                         o_id = VORIZT1259,
                         d_id = VDESZT1259,
                         pop = workers_transit,
                         opp = jobs_transit,
                         r = catch,
                         f = f_transit),
         SA_ij_transit = ifelse(is.nan(SA_ij_transit), 0, SA_ij_transit),
         SA_ij_bike = sp_avail(., 
                         o_id = VORIZT1259,
                         d_id = VDESZT1259,
                         pop = workers_bike,
                         opp = jobs_bike,
                         r = catch,
                         f = f_bike),
         SA_ij_bike = ifelse(is.nan(SA_ij_bike), 0, SA_ij_bike),
         SA_ij_walk = sp_avail(., 
                         o_id = VORIZT1259,
                         d_id = VDESZT1259,
                         pop = workers_walk,
                         opp = jobs_walk,
                         r = catch,
                         f = f_walk),
         SA_ij_walk = ifelse(is.nan(SA_ij_walk), 0, SA_ij_walk))

#verify that the sum of all jobs is consistent with the number of jobs
sum(OD_SA$SA_ij_car) + sum(OD_SA$SA_ij_transit) + sum(OD_SA$SA_ij_bike) + sum(OD_SA$SA_ij_walk)

#verify that the sum of all jobs is consistent with the number of jobs - from another perspective
sum_jobs <- OD_SA %>% 
  group_by(VDESZT1259) %>% 
  summarise(jobs = mean(jobs_transit) + mean(jobs_car)  + mean(jobs_walk) + mean(jobs_bike))
sum(sum_jobs$jobs, na.rm = T)

# same number excellent! there are 27434 jobs total and there are 27795 workers (and subsequently 27795 trips total). The mismatch in numbers arises because we only survey people who's work trip originates in the Madrid region (from home). This means that workers who travel outside of the madrid region still count as a worker (and a trip) + increase 'competition' for the zone.... maybe we don't want this?

```

Let's try calculating SA from a different approch:
```{r}
# Also another question - this configuration assumes competition is contained within modes - i.e., transit workers compete for transit jobs, etc. This isn't the most realistic I think. So let's sum up all the mode specific workers and jobs and impedance values BEFORE calculating SA so we only have 1 SA

#add new summed columns
OD_SA <- OD_SA %>%
  mutate(workers_ = (workers_car+ workers_transit+ workers_bike+ workers_walk),
         workers_ = ifelse(is.na(workers_), 0, workers_),
         jobs_ = (jobs_car + jobs_transit + jobs_bike + jobs_walk),
         jobs_ = ifelse(is.na(jobs_), 0, jobs_),
         f_ = f_car + f_transit + f_bike + f_walk)

#calculate spatial availability
OD_SA2 <- OD_SA %>%
  mutate(catch = 1) %>%
  mutate(SA_ij = sp_avail(., 
                         o_id = VORIZT1259,
                         d_id = VDESZT1259,
                         pop = workers_,
                         opp = jobs_,
                         r = catch,
                         f = f_),
         SA_ij = ifelse(is.nan(SA_ij), 0, SA_ij))

#verify that the sum of all jobs is consistent with the number of jobs
sum(OD_SA2$SA_ij)

#verify that the sum of all jobs is consistent with the number of jobs - from another perspective
sum_jobs <- OD_SA2 %>% 
  group_by(VDESZT1259) %>% 
  summarise(jobs = mean(jobs_))
sum(sum_jobs$jobs, na.rm = T)
#Perfect it works! 
```

Let's aggregate the spatial availability by origin zone for these 2 methods (calculating SA separately... and calculating 1 SA that combines all modes)
```{r}
#aggregating spatial availability  
MAD_availability <- OD_SA2 %>%
  group_by(VORIZT1259) %>%
  summarize(SA_i2 = sum(SA_ij),
            SA_i1 = sum(SA_ij_car) + sum(SA_ij_bike) + sum(SA_ij_transit) + sum(SA_ij_walk),
            workers_ = mean(workers_car, na.rm=T) + mean(workers_transit, na.rm=T) + mean(workers_bike, na.rm=T) + mean(workers_walk, na.rm=T),
            avgtt_i = mean(mean(tt_transit) + mean(tt_walk) + mean(tt_car) + mean(tt_bike)),
            avg_f_i = mean(mean(f_transit) + mean(f_walk) + mean(f_car) + mean(f_bike)))

#Merge TO availability calculation to the TAZ sf object created for accessibility above:
zn_test <- zn %>% merge(MAD_availability, by.x=c("ZT1259"), by.y=c("VORIZT1259"), all.x=T) 
```

```{r create-bounding-boxes}
#create the bounding box for the Toronto region + a little extra
bbox_new1 <- st_bbox(zn_test%>%filter(Municipality == "Madrid")) # current bounding box

xrange <- bbox_new1$xmax - bbox_new1$xmin # range of x values
yrange <- bbox_new1$ymax - bbox_new1$ymin # range of y values

bbox_new1[1] <- bbox_new1[1] - (0.0 * xrange) # xmin - left
bbox_new1[3] <- bbox_new1[3] + (0.0 * xrange) # xmax - right
bbox_new1[2] <- bbox_new1[2] - (0.0 * yrange) # ymin - bottom
bbox_new1[4] <- bbox_new1[4] + (0.0 * yrange) # ymax - top

bbox_Madrid_muni <- bbox_new1 %>%  # take the bounding box ...
  st_as_sfc()

Madrid_muni_border <- zn_test%>%filter(Municipality == "Madrid")%>%st_union()
```

plot 1 (SA for each mode calculated seperated and then added together) and plot 2 (SA for all modes calculated together):
```{r}
#plot 1 main plot
plot_MAD_avail1 <- ggplot() +
  geom_sf(data = zn_test, aes(fill= SA_i1), color = NA) + #data
    scale_fill_distiller(palette = "RdYlGn", #legend scale bar
                         name = expression(V["i"]),
                         na.value = "grey90",
                         limits = c( round(min(zn_test$SA_i1, na.rm=T)), round(max(zn_test$SA_i1, na.rm=T)))) +
  geom_sf(data = Madrid_muni_border, # border for Madrid muni
          colour=alpha("dimgrey",1), 
          size = 0.5, fill=NA, 
          show.legend = "polygon") +
  annotation_north_arrow(location = "tl", # north arrow for both the main plot and inset
                         height = unit(0.8, "cm"), 
                         width = unit(0.8, "cm"),
                         style = north_arrow_orienteering(line_width = 0.25,
                                                          line_col = "dimgrey", 
                                                          fill = c("white","dimgrey"))) +
  annotation_scale(bar_cols = c("dimgrey", "white"), # scale bar for both the main plot and inset
                   height = unit(0.15, "cm")) +
  theme_void() +
  theme(legend.position = c(1.10, 0.25), 
        plot.margin = margin(t = 0, r = 0, b = 0, l = -80)) # positing legend and clipping out white space

# plot 1 + inset map
plot_MAD_avail1 <- ggdraw(plot_MAD_avail1) +
  draw_plot({plot_MAD_avail1 + coord_sf(xlim = st_coordinates(bbox_Madrid_muni)[c(1,2),1],
                                         ylim = st_coordinates(bbox_Madrid_muni)[c(2,3),2],
                                         expand = FALSE) +
      theme(legend.position = "none")},
    x = 0.65, y = 0.50,  # The distance along a (0,1) x-axis (and y axis) to draw the left edge (the bottom edge) of the plot
    width = 0.46, height = 0.46) # The width and height of the plot expressed as proportion of the entire ggdraw object

#plot 2
plot_MAD_avail2 <- ggplot() +
  geom_sf(data = zn_test, aes(fill= SA_i2), color = NA) + #data
    scale_fill_distiller(palette = "RdYlGn", #legend scale bar
                         name = expression(V["i"]),
                         na.value = "grey90",
                         limits = c( round(min(zn_test$SA_i2, na.rm=T)), round(max(zn_test$SA_i2, na.rm=T)))) +
  geom_sf(data = Madrid_muni_border, # border for Madrid muni
          colour=alpha("dimgrey",1), 
          size = 0.5, fill=NA, 
          show.legend = "polygon") +
  annotation_north_arrow(location = "tl", # north arrow for both the main plot and inset
                         height = unit(0.8, "cm"), 
                         width = unit(0.8, "cm"),
                         style = north_arrow_orienteering(line_width = 0.25,
                                                          line_col = "dimgrey", 
                                                          fill = c("white","dimgrey"))) +
  annotation_scale(bar_cols = c("dimgrey", "white"), # scale bar for both the main plot and inset
                   height = unit(0.15, "cm")) +
  theme_void() +
  theme(legend.position = c(1.10, 0.25), 
        plot.margin = margin(t = 0, r = 0, b = 0, l = -80)) # positing legend and clipping out white space

# plot 1 + inset map
plot_MAD_avail2 <- ggdraw(plot_MAD_avail2) +
  draw_plot({plot_MAD_avail2 + coord_sf(xlim = st_coordinates(bbox_Madrid_muni)[c(1,2),1],
                                         ylim = st_coordinates(bbox_Madrid_muni)[c(2,3),2],
                                         expand = FALSE) +
      theme(legend.position = "none")},
    x = 0.65, y = 0.50,  # The distance along a (0,1) x-axis (and y axis) to draw the left edge (the bottom edge) of the plot
    width = 0.46, height = 0.46) # The width and height of the plot expressed as proportion of the entire ggdraw object

plot_MAD_avail1
plot_MAD_avail2
```
We can normalize it by the number of workers so the numeric value corresponds to available jobs per worker. First plot 1, then plot 2.
```{r}
#plot 1 main plot
plot_MAD_avail_perworker_1 <- ggplot() +
  geom_sf(data = zn_test, aes(fill= SA_i1/workers_), color = NA) + #data
    scale_fill_distiller(palette = "RdYlGn", #legend scale bar
                         name = expression(v["i"]),
                         na.value = "grey90",
                         limits = c( round(min(zn_test$SA_i1/zn_test$workers_, na.rm=T)), 
                                     round(max(zn_test$SA_i1/zn_test$workers_, na.rm=T))))+
  geom_sf(data = Madrid_muni_border, # border for Madrid muni
          colour=alpha("dimgrey",1), 
          size = 0.5, fill=NA, 
          show.legend = "polygon") +
  annotation_north_arrow(location = "tl", # north arrow for both the main plot and inset
                         height = unit(0.8, "cm"), 
                         width = unit(0.8, "cm"),
                         style = north_arrow_orienteering(line_width = 0.25,
                                                          line_col = "dimgrey", 
                                                          fill = c("white","dimgrey"))) +
  annotation_scale(bar_cols = c("dimgrey", "white"), # scale bar for both the main plot and inset
                   height = unit(0.15, "cm")) +
  theme_void() +
  theme(legend.position = c(1.10, 0.25), 
        plot.margin = margin(t = 0, r = 0, b = 0, l = -80)) # positing legend and clipping out white space

# plot 1 + inset map
plot_MAD_avail_perworker_1 <- ggdraw(plot_MAD_avail_perworker_1) +
  draw_plot({plot_MAD_avail_perworker_1 + coord_sf(xlim = st_coordinates(bbox_Madrid_muni)[c(1,2),1],
                                         ylim = st_coordinates(bbox_Madrid_muni)[c(2,3),2],
                                         expand = FALSE) +
      theme(legend.position = "none")},
    x = 0.65, y = 0.50,  # The distance along a (0,1) x-axis (and y axis) to draw the left edge (the bottom edge) of the plot
    width = 0.46, height = 0.46) # The width and height of the plot expressed as proportion of the entire ggdraw object

#plot 2
plot_MAD_avail_perworker_2 <- ggplot() +
  geom_sf(data = zn_test, aes(fill= SA_i2/workers_), color = NA) + #data
    scale_fill_distiller(palette = "RdYlGn", #legend scale bar
                         name = expression(v["i"]),
                         na.value = "grey90",
                         limits = c( round(min(zn_test$SA_i2/zn_test$workers_, na.rm=T)), 
                                     round(max(zn_test$SA_i2/zn_test$workers_, na.rm=T))))+
  geom_sf(data = Madrid_muni_border, # border for Madrid muni
          colour=alpha("dimgrey",1), 
          size = 0.5, fill=NA, 
          show.legend = "polygon") +
  annotation_north_arrow(location = "tl", # north arrow for both the main plot and inset
                         height = unit(0.8, "cm"), 
                         width = unit(0.8, "cm"),
                         style = north_arrow_orienteering(line_width = 0.25,
                                                          line_col = "dimgrey", 
                                                          fill = c("white","dimgrey"))) +
  annotation_scale(bar_cols = c("dimgrey", "white"), # scale bar for both the main plot and inset
                   height = unit(0.15, "cm")) +
  theme_void() +
  theme(legend.position = c(1.10, 0.25), 
        plot.margin = margin(t = 0, r = 0, b = 0, l = -80)) # positing legend and clipping out white space

# plot 1 + inset map
plot_MAD_avail_perworker_2 <- ggdraw(plot_MAD_avail_perworker_2) +
  draw_plot({plot_MAD_avail_perworker_2 + coord_sf(xlim = st_coordinates(bbox_Madrid_muni)[c(1,2),1],
                                         ylim = st_coordinates(bbox_Madrid_muni)[c(2,3),2],
                                         expand = FALSE) +
      theme(legend.position = "none")},
    x = 0.65, y = 0.50,  # The distance along a (0,1) x-axis (and y axis) to draw the left edge (the bottom edge) of the plot
    width = 0.46, height = 0.46) # The width and height of the plot expressed as proportion of the entire ggdraw object

plot_MAD_avail_perworker_1
plot_MAD_avail_perworker_2
```

## An optimistic scenario - no cars going into madrid municipality

Let's test out a scenario - all car trips going into madrid municipality are now transit trips. How will this impact the SA profile? We will using the 2nd method (calculating SA within all modes).

First, let's re-assign to-madrid muni car trips to transit trips:
```{r}
Madrid_muni <- zn %>% filter(Municipality == "Madrid") %>% 
  dplyr::transmute(ZT1259,
                   MM = 1)

OD_v_HW_optimistic <- OD_v_HW  %>% merge(Madrid_muni, by.x=("VDESZT1259"), by.y=("ZT1259"), all.x=T)

#reassigning car trips to transit
OD_v_HW_optimistic1 <- OD_v_HW_optimistic  %>% filter(MM == 1) %>% rowwise() %>%
  mutate(trips_transit = sum(trips_transit,trips_car_motor, na.rm = T),
         trips_car_motor = 0)

OD_v_HW_optimistic2 <- OD_v_HW_optimistic  %>% filter(is.na(MM))

OD_v_HW_optimistic <- rbind(OD_v_HW_optimistic1,OD_v_HW_optimistic2)

workers <- OD_v_HW_optimistic %>%
  group_by(VORIZT1259) %>%
  summarize(workers_transit = sum(trips_transit),
            workers_car = sum(trips_car_motor),
            workers_bike = sum(trips_bike),
            workers_walk = sum(trips_walk),
            .groups = "drop") %>%
  transmute(ZT1259 = VORIZT1259,
            workers_transit,
            workers_car,
            workers_bike,
            workers_walk)

jobs <- OD_v_HW_optimistic %>%
  group_by(VDESZT1259) %>%
  summarize(jobs_transit = sum(trips_transit),
            jobs_car = sum(trips_car_motor),
            jobs_bike = sum(trips_bike),
            jobs_walk = sum(trips_walk),
            .groups = "drop") %>%
  transmute(ZT1259 = VDESZT1259,
            jobs_transit,
            jobs_car,
            jobs_bike,
            jobs_walk)
```

Now again, join the workers and jobs to the ZN simple features object:
```{r}
load("data/zn.RData")

zn <- zn %>%
  # Join workers
  left_join(workers,
            by = "ZT1259") %>%
  # Join jobs
  left_join(jobs,
            by = "ZT1259") %>%
  # Replace NAs with zeros
  mutate(workers_transit = tidyr::replace_na(workers_transit, 0),
         workers_car = tidyr::replace_na(workers_car, 0),
         workers_bike = tidyr::replace_na(workers_bike, 0),
         workers_walk = tidyr::replace_na(workers_walk, 0),
         jobs_transit = tidyr::replace_na(jobs_transit, 0),
         jobs_car = tidyr::replace_na(jobs_car, 0),
         jobs_bike = tidyr::replace_na(jobs_bike, 0),
         jobs_walk = tidyr::replace_na(jobs_walk, 0))
```

add the number of workers and jobs from each zona to each trip:
```{r}
#add the number of jobs and workers to the od_ft_tt matrix
OD_optimistic <- OD_v_HW_optimistic %>% merge(zn %>% dplyr::select(ZT1259, workers_transit,workers_car,workers_bike,workers_walk) %>% st_drop_geometry(),
                   by.x = "VORIZT1259", by.y="ZT1259", all.x = TRUE)

OD_optimistic <- OD_optimistic %>% merge(zn %>% dplyr::select(ZT1259, jobs_transit,jobs_car,jobs_bike,jobs_walk) %>% st_drop_geometry(),
                   by.x = "VDESZT1259", by.y="ZT1259", all.x = TRUE)

OD_optimistic <- OD_optimistic %>% 
  mutate(workers_transit = ifelse(is.na(workers_transit), 0, workers_transit),
         workers_car = ifelse(is.na(workers_car), 0, workers_car),
         workers_bike = ifelse(is.na(workers_bike), 0, workers_bike),
         workers_walk = ifelse(is.na(workers_walk), 0, workers_walk),
         jobs_car = ifelse(is.na(jobs_car), 0, jobs_car),
         jobs_transit = ifelse(is.na(jobs_transit), 0, jobs_transit),
         jobs_bike = ifelse(is.na(jobs_bike), 0, jobs_bike),
         jobs_walk = ifelse(is.na(jobs_walk), 0, jobs_walk))
```

now reshape the data, each row will represent an any origin to to 1 destination (i.e., destination row is unique). We will average the travel times, by mode, and then use this average travel time to find the 'f' (impedance value for travel for each mode):
```{r}
#consolidate trips so each trip is unique. workers = jobs.
OD_optimistic <- OD_optimistic %>%
  group_by(VORIZT1259,VDESZT1259) %>%
  summarise(trips_transit = sum(trips_transit),
            trips_car_motor = sum(trips_car_motor),
            trips_bike = sum(trips_bike),
            trips_walk = sum(trips_walk),
            workers_transit = mean(workers_transit),
            # workers_transit = ifelse(is.na(workers_transit), 0, workers_transit),
            workers_car = mean(workers_car),
            # workers_car = ifelse(is.na(workers_car), 0, workers_car),
            workers_bike = mean(workers_bike),
            # workers_bike = ifelse(is.na(workers_bike), 0, workers_bike),
            workers_walk = mean(workers_walk),
            # workers_walk = ifelse(is.na(workers_walk), 0, workers_walk),
            jobs_transit = mean(jobs_transit),
            jobs_car = mean(jobs_car),
            jobs_bike = mean(jobs_bike),
            jobs_walk = mean(jobs_walk),
            # jobs_car = ifelse(is.na(jobs_car), 0, jobs_car),
            # jobs_transit = ifelse(is.na(jobs_transit), 0, jobs_transit),
            # jobs_bike = ifelse(is.na(jobs_bike), 0, jobs_bike),
            # jobs_walk = ifelse(is.na(jobs_walk), 0, jobs_walk),
            tt_transit = ifelse(trips_transit != 0, mean(TIEMPO_VIAJES, na.rm=T), NA),
            tt_car = ifelse(trips_car_motor != 0, mean(TIEMPO_VIAJES, na.rm=T), NA),
            tt_bike = ifelse(trips_bike != 0, mean(TIEMPO_VIAJES, na.rm=T), NA),
            tt_walk = ifelse(trips_walk != 0, mean(TIEMPO_VIAJES, na.rm=T), NA)) %>% ungroup()

```

now estimate f value using these averaged travel times:
```{r, echo=FALSE, warning=FALSE, message=FALSE}
#dgamma is the density plot for gamma function and dlnorm is the desntiy plot for the lognormal function; adding values as "f" our impedance function
OD_optimistic <- OD_optimistic %>%
  mutate(f_car = dgamma(tt_car, HW_car_motor_gamma$estimate["shape"], HW_car_motor_gamma$estimate["rate"]),
         f_car = ifelse(is.na(f_car), 0, f_car),
         f_transit = dlnorm(tt_transit, HW_transit_lnorm$estimate["meanlog"], HW_transit_lnorm$estimate["sdlog"]),
         f_transit = ifelse(is.na(f_transit), 0, f_transit),
         f_bike = dlnorm(tt_bike, HW_bike_lnorm$estimate["meanlog"], HW_bike_lnorm$estimate["sdlog"]),
         f_bike = ifelse(is.na(f_bike), 0, f_bike),
         f_walk = dlnorm(tt_walk, HW_walk_lnorm$estimate["meanlog"], HW_walk_lnorm$estimate["sdlog"]),
         f_walk = ifelse(is.na(f_walk), 0, f_walk))
```

Calculating SA:
```{r}
#add new summed columns
OD_optimistic <- OD_optimistic %>%
  mutate(workers_ = (workers_car+ workers_transit+ workers_bike+ workers_walk),
         workers_ = ifelse(is.na(workers_), 0, workers_),
         jobs_ = (jobs_car + jobs_transit + jobs_bike + jobs_walk),
         jobs_ = ifelse(is.na(jobs_), 0, jobs_),
         f_ = f_car + f_transit + f_bike + f_walk)

#calculate spatial availability
OD_optimistic <- OD_optimistic %>%
  mutate(catch = 1) %>%
  mutate(SA_ij = sp_avail(., 
                         o_id = VORIZT1259,
                         d_id = VDESZT1259,
                         pop = workers_,
                         opp = jobs_,
                         r = catch,
                         f = f_),
         SA_ij = ifelse(is.nan(SA_ij), 0, SA_ij))

#verify that the sum of all jobs is consistent with the number of jobs
sum(OD_optimistic$SA_ij)

#verify that the sum of all jobs is consistent with the number of jobs - from another perspective
sum_jobs <- OD_optimistic %>% 
  group_by(VDESZT1259) %>% 
  summarise(jobs = mean(jobs_))
sum(sum_jobs$jobs, na.rm = T)
#Perfect it works! 
```

Let's aggregate the spatial availability by origin zone:
```{r}
#aggregating spatial availability  
MAD_availability <- OD_optimistic %>%
  group_by(VORIZT1259) %>%
  summarize(SA_i3 = sum(SA_ij),
            workers_ = mean(workers_car, na.rm=T) + mean(workers_transit, na.rm=T) + mean(workers_bike, na.rm=T) + mean(workers_walk, na.rm=T),
            avgtt_i = mean(mean(tt_transit) + mean(tt_walk) + mean(tt_car) + mean(tt_bike)),
            avg_f_i = mean(mean(f_transit) + mean(f_walk) + mean(f_car) + mean(f_bike)))

#Merge TO availability calculation to the TAZ sf object created for accessibility above:
zn_test3 <- zn %>% merge(MAD_availability, by.x=c("ZT1259"), by.y=c("VORIZT1259"), all.x=T) 
```

Here we plot plot 2 (the original scenario) and plot 3 ( this optimistic scenario)
```{r}
#plot 2
plot_MAD_avail_perworker_3 <- ggplot() +
  geom_sf(data = zn_test3, aes(fill= SA_i3/workers_), color = NA) + #data
    scale_fill_distiller(palette = "RdYlGn", #legend scale bar
                         name = expression(v["i"]),
                         na.value = "grey90",
                         limits = c( round(min(zn_test3$SA_i3/zn_test3$workers_, na.rm=T)), 
                                     round(max(zn_test3$SA_i3/zn_test3$workers_, na.rm=T))))+
  geom_sf(data = Madrid_muni_border, # border for Madrid muni
          colour=alpha("dimgrey",1), 
          size = 0.5, fill=NA, 
          show.legend = "polygon") +
  annotation_north_arrow(location = "tl", # north arrow for both the main plot and inset
                         height = unit(0.8, "cm"), 
                         width = unit(0.8, "cm"),
                         style = north_arrow_orienteering(line_width = 0.25,
                                                          line_col = "dimgrey", 
                                                          fill = c("white","dimgrey"))) +
  annotation_scale(bar_cols = c("dimgrey", "white"), # scale bar for both the main plot and inset
                   height = unit(0.15, "cm")) +
  theme_void() +
  theme(legend.position = c(1.10, 0.25), 
        plot.margin = margin(t = 0, r = 0, b = 0, l = -80)) # positing legend and clipping out white space

# plot 1 + inset map
plot_MAD_avail_perworker_3 <- ggdraw(plot_MAD_avail_perworker_3) +
  draw_plot({plot_MAD_avail_perworker_3 + coord_sf(xlim = st_coordinates(bbox_Madrid_muni)[c(1,2),1],
                                         ylim = st_coordinates(bbox_Madrid_muni)[c(2,3),2],
                                         expand = FALSE) +
      theme(legend.position = "none")},
    x = 0.65, y = 0.50,  # The distance along a (0,1) x-axis (and y axis) to draw the left edge (the bottom edge) of the plot
    width = 0.46, height = 0.46) # The width and height of the plot expressed as proportion of the entire ggdraw object

plot_MAD_avail_perworker_2
plot_MAD_avail_perworker_3
```

As can be seen, the change (visability) is not that dramatic.