### 2023-08-25 Journal-manuscript/Title-page.Rmd
Added title page as requested by editor

---
title: "Multimodal spatial availability: a singly-constrained measure of competitive accessibility considering multiple modes"
author:
  - name: Anastasia Soukhov
    corresponding: true
    position: PhD Candidate
    email: soukhoa@mcmaster.ca (AS)
    affiliation: McMaster
  - name: Javier Tarriño-Ortiz
    position: PhD Graduate
    email: javier.tarrino.ortiz@upm.es
    affiliation: Madrid
  - name: Julio A. Soria-Lara
    position: Associate Professor
    email: julio.soria-lara@upm.es
    affiliation: Madrid
  - name: Antonio Páez
    position: Associate Professor
    email: paezha@mcmaster.ca
    affiliation: McMaster
address:
  - code: McMaster
    address: Department of Earth, Environment and Society, McMaster University, Hamilton, Canada
  - code: Madrid
    address: Centro de Investigación del Transporte (TRANSyT), Universidad Politécnica de Madrid, Madrid, Spain

output: rticles::plos_article
header-includes:
  - \usepackage{booktabs} 
  - \usepackage{sectsty} \sectionfont{\centering}
  - \renewcommand{\contentsname}{}\vspace{-2cm} 
  - \usepackage{pdfpages}2
# plos.csl can be download and used locally
csl: http://www.zotero.org/styles/plos
---

\includepdf[pages={-}]{Journal-manuscript.pdf}

\newpage