### 2023-11-09 README.Rmd
Updated README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Multimodal spatial availability: a singly-constrained measure of accessibility considering multiple modes

<!-- badges: start -->
<!-- badges: end -->

Anastasia Soukhov (McMaster University)
Javier Tarriño-Ortiz (Universidad Politécnica de Madrid)
Julio Soria-Lara (Universidad Politécnica de Madrid)
Antonio Páez (McMaster University)

Recent research has aimed to address the way opportunities are counted in accessibility analysis. In conventional accessibility measures, opportunities are often multiply counted, which leads to values of accessibility that are difficult to interpret. Constraining the calculations to match a known quantity ensures that the measurements sum up to a predetermined quantity (i.e., the total number of opportunities), and so each value can be meaningfully related to this total. A recent such effort is spatial availability, a singly-constrained accessibility measure. In this paper we extend spatial availability for use in the case of multiple modes or, more generally, heterogeneous population segments with distinct travel behaviors. After deriving a multimodal version of spatial availability We proceed to illustrate its features using a synthetic example. Next, we apply it to an empirical example in Madrid. We conclude the paper with suggestions for future research.

Submitted to PLoS ONE.